Optimal (Multiway) Spatial Joins

### 2024-06-02 msj.tex
yufei

%In other words, both $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ must have size at least 1.

@@

To compute $R^*_1, R^*_2, ..., R^*_{k-2}$, we start by deriving $\maxtop(r')$ for each rectangle $r'$ in $R'_1 \cup  ... \cup R'_{k-2}$. For this purpose, first initialize $\maxtop(r') = -\infty$ for each such $r'$. Recall that $r'$ is the trimmed rectangle of some rectangle $r$ in $R_1 \cup ... \cup R_{k-2}$. We store $\maxtop(r')$ in the array cell allocated to $r$.

Then, we scan $\J(R_1',...,R_{k-2}')$. For each tuple $\bm{t}$ therein, use $O(k)$ time to identify the rectangle $r' = \gbot(\bm{t})$, and then update in constant time $\maxtop(r')$ to the maximum between its current value and $\ytop(r')$. The scan requires $O(n + k \cdot \out)$ time.

@@

% We will provide a way to map each pair $(\bm{t'}, r)$ satisfyint that $\bm{t'}\in \J(R_1,...,R_{k-3},H',V)$ and $r \in \contained_{R_{k-2}}(\bm{t})$ to a unique tuple in $\J_2$. Consider an arbitrary such pair $(\bm{t'}, r)$. Let $h$ be the full segment of $\bm{t'}[k-2]$. We map $(\bm{t'}, r)$ to the tuple $\bm{t} = (\bm{t}'[1],...,\bm{t}'[k-3], r, h, \bm{t}'[k-1])$. By Statement (1) and the fact that $r\in \contained_{R_{k-2}}(\bm{t'})$, the $k$-tuple $\bm{t}$ belongs to $\J_2$.
%
% \vgap
%
% It remains to prove that no two distinct pair $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ will be mapped to the same tuple in $\J_2$. Assume, on the contrary, $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ are mapped to the same tuple $\bm{t} \in \J_2$. Then, $r_1 = \bm{t}[k-2] = r_2$, and $\bm{t}_1[i] = \bm{t}[i] = \bm{t}_2[i]$ for $i \in [k-3]$. Moreover, both $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are the trimmed segment of $\bm{t}[k-1]$. Since the trimmed segment is unique, $\bm{t}_1[k-2] = \bm{t}_2[k-2]$. Hence, $\bm{t}_1 = \bm{t}_2$, which contradicts the assumption. Therefore, we have $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the second inequality is due to Definition \eqref{eqn:hv:type2:J2}.

\extraspacing {\bf Computing $\bm{H^*}$.} W.l.o.g., we assume that each segment in the input $H$ is given a distinct integer ID in $[|H|]$. This allows us to create an array of size $|H| < n$ and allocate an array cell to each $h \in H$, with the property that the cell can be accessed by the ID of $h$ in constant time.

@@

To compute $H^*$, we start by deriving $\minleft(h')$ for each segment $h'$ in $H'$. For this purpose, first initialize $\minleft(h') = \infty$ for each such $h'$. Recall that $h'$ is the trimmed segment of some segment $h$ in $H$. We store $\minleft(h')$ in the array cell allocated to $h$.

Then, we scan $\J(R_1,...,R_{k-3},H',V)$. For each tuple $\bm{t}$ therein, update in constant time $\minleft(h')$ to the minimum between its current value and the x-coordinate of $\bm{t}[k-1]$. The scan requires $O(n + k \cdot \out)$ time.

@@

\input{5.3.tex}

>>>>

@@

To compute $R^*_1, R^*_2, ..., R^*_{k-2}$, we start by deriving $\maxtop(r')$ for each rectangle $r'$ in $R'_1 \cup  ... \cup R'_{k-2}$. For this purpose, first initialize $\maxtop(r') = -\infty$ for each such $r'$. Recall that $r'$ is the trimmed rectangle of some rectangle $r$ in $R_1 \cup ... \cup R_{k-2}$. We store $\maxtop(r')$ in the array cell allocated to $r$. Then, we scan $\J(R_1',...,R_{k-2}')$. For each tuple $\bm{t}$ therein, use $O(k)$ time to identify the rectangle $r' = \gbot(\bm{t})$, and then update in constant time $\maxtop(r')$ to the maximum between its current value and $\ytop(r')$. The scan requires $O(n + k \cdot \out)$ time.

@@

\extraspacing {\bf Computing $\bm{H^*}$.} We assume, w.l.o.g., that each segment in the input $H$ is given a distinct integer ID in $[|H|]$. This allows us to create an array of size $|H| < n$ and allocate an array cell to each $h \in H$, with the property that the cell can be accessed by the ID of $h$ in constant time.

@@

To compute $H^*$, we start by deriving $\minleft(h')$ for each segment $h'$ in $H'$. For this purpose, first initialize $\minleft(h') = \infty$ for each such $h'$. Recall that $h'$ is the trimmed segment of some segment $h$ in $H$. We store $\minleft(h')$ in the array cell allocated to $h$. Then, we scan $\J(R_1,...,R_{k-3},H',V)$. For each tuple $\bm{t}$ therein, update in constant time $\minleft(h')$ to the minimum between its current value and the x-coordinate of $\bm{t}[k-1]$. The scan requires $O(n + k \cdot \out)$ time.

@@

\noindent \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\cross_{R_{k-2}}(\bm{t}) = \cross_{R_{k-2}}(h^*)$. We can then derive that
\myeqn{
    \sum_{h^* \in H^*} |\cross_{R_{k-2}}(h^*)| \le \sum_{\bm{t} \in \J(R_1,...,R_{k-3},H',V)} |\cross_{R_{k-2}}(\bm{t})| \leq \out, \nn
} where the last inequality is due to Statement (3) of Lemma~\ref{lmm:hv:type2:properties}.

Now, we explain the mapping. Consider an arbitrary $h^* \in H^*$. Recall that $h^*$ is the minimal segment of some segment $h' \in H'$. Specifically, if $h' = [x_1,x_2]\times y$, then $h^* = [x_1,\minleft(h')]\times y$. By the definition of $\minleft(h')$ in \eqref{eqn:hv:type2:H*}, there exists a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\bm{t}[k-2] = h'$ and $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$}$. We map $h^*$ to $\bm{t}$.

Next, we will prove that a rectangle $r_{k-2} \in \cross_{R_{k-2}}(h^*)$ if and only if $r_{k-2} \in \cross_{R_{k-2}}(\bm{t})$. By the definition of $\cross_{R_{k-2}}(h^*)$ and $\cross_{R_{k-2}}(\bm{t})$ (see Definition~\eqref{eqn:contained} and \eqref{eqn:type2:contained-t}), it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$. Let $x$ be the $x$-coordinate of the vertical segment $\bm{t}[k-1]$. Then, $B_\bm{t}$, which is the intersection of $h'$ and $v$, is $(x,y)$, where $y$ is the $y$-coordinate of $h'$. The effective horizontal segment of $\bm{t}$ is therefore $[x_1,x]\times y$. Recall that $h^* = [x_1,\minleft(h')]\times y$, where $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$} = x$. Therefore, $h^*$ is the same as the effective horizontal segment of $\bm{t}$.

It remains to show that no two distinct rectangles $h_1^*, h_2^* \in H^*$ can be mapped to the same tuple in $\J(R_1,...,R_{k-3}, H',V)$. Assume, on the contrary, that $h_1^*$ and $h_2^*$ are mapped to the same tuple $\bm{t}\in \J(R_1,...,R_{k-3}, H',V)$. Suppose that $h_1^*$ (resp., $h_2^*$) is the minimal segment of $r_1'$ (resp., $r_2'$). Under our mapping, it must be true that  $h_1' = h_2' = \bm{t}[k-2]$. However, the distinctness of $h_1^*$ and $h_2^*$ requires $h_1' \ne h_2'$, which yields a contradiction.

\extraspacing \underline{\em Proof of Statement (2).}
Let $h', h^*$, and $\bm{t}$ be as defined in statement (2). Let us represent $h'$ as $[x_1,x_2] \times y$. Accordingly, $h^*$ can be written as $[x_1, \minleft(h')] \times y$. Additionally, let $x$ be the x-coordinate of the vertical segment $\bm{t}[k-1]$. Then, $B_\bm{t}$, which is the intersection of $h'$ and $v$, is $(x,y)$. The effective horizontal segment of $\bm{t}$ can be denoted as $[x_1,x] \times y$.

We will first prove $\contained_{R_{k-2}}(\bm{t})\subseteq \cross_{R_{k-2}}(h^*)$. For this purpose, we will show that any rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$ also covers $h^*$. It suffices to show that $h^*$ is contained in the effective horizontal segment of $\bm{t}$. Combining the definition of $\minleft(h')$ (see \eqref{eqn:minleft}), and the fact that $x$ is the x-coordinate of $\bm{t}[k-1]$, where $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, we have $\minleft(h') \le x$. This implies that $h^* = [x_1, \minleft(h')] \times y$ is contained in $[x_1, x] \times y$, i.e. the effective horizontal segment of $\bm{t}$.

Next, assuming that the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ are sorted in descending order of $\xright(r)$, then we will prove that $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order. It suffices to establish the following equivalent statement:

\minipg{0.8\linewidth} {
    If a rectangle $r \in \contained_{R_{k-2}}(h^*)$ satisfies that $\xright(r)$ is no less than the x-coordinate of $\bm{t}[k-1]$, then $r$ must be in $\contained_{R_{k-2}}(\bm{t})$.
}

\noindent To prove the above, we must explain why $r$ contains the effective horizontal segment of $\bm{t}$. Since $r \in \contained_{R_{k-2}}(h^*)$, $r$ covers the segment $[x_1, \minleft(h')] \times y$. Therefore, $[x_1, \minleft(h')] \subseteq [\xleft(r), \xright(r)]$ and $y \in [\ybot(r), \ytop(r)]$.
As $\xright(r)$ is no less than the x-coordinate of $\bm{t}[k-1]$, which is denoted as $x$, we have $[x_1, x] \subseteq [\xleft(r), \xright(r)]$. This implies that $[x_1, x] \times y$, i.e. the effective horizontal segment of $\bm{t}$, is contained in $r$.
